Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-12-02 4.tex
First complete draft!

It is a curious fact that many constraints on underlying representations are elegantly described using the units of prosody.
In some cases, these are part of the system of contrast, such as long segments in moraic theory \citep[e.g.,][]{Davis1999}. For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants. However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:

\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
\label{mcl}
A medial cluster must consist of a well-formed medial coda and a well-formed medial onset
\end{example}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{Dorsal-labial cluster attestation in the lexical sample}
\label{dltab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{ABA cluster attestation in the lexical sample}
\label{abatab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
\label{ovatab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.4\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{Nasal place assimilation cluster attestation in the lexical sample}
\label{npatab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{Degemination in the lexical sample}
\label{degemtab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
\label{cmresults}
\end{table}

@@

\begin{figure}[ht]
\centering
\includegraphics{co.pdf}
\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{cao}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line). Simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{sim}
\end{figure}

>>>>

It is a curious fact that many constraints on underlying representations are elegantly described using the units of prosody.
In some cases, these are part of the system of contrast, such as long segments in moraic theory \citep[e.g.,][]{Davis1999}. For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants. 
However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:

\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
\label{mcl}
A medial cluster consists maximally of a well-formed medial coda and a well-formed medial onset
\end{example}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{Dorsal-labial cluster attestation in the lexical sample}
\label{dltab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{ABA cluster attestation in the lexical sample}
\label{abatab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
\label{ovatab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.4\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{Nasal place assimilation cluster attestation in the lexical sample}
\label{npatab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{Degemination in the lexical sample}
\label{degemtab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task}
\label{cmresults}
\end{table}
%; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}

@@

\begin{figure}[t]
\centering
\includegraphics{co.pdf}
\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
\label{cao}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line); simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
\label{sim}
\end{figure}
